---
title: "Figure 2 and related supplement tables"
author: "Meng-Ching Ko (MaggieMCKO)"
date: "07/17/2024"
output:
  html_document:
    df_print: paged
  pdf_document: default
editor_options: 
  chunk_output_type: console
---

This [R Markdown](http://rmarkdown.rstudio.com) Notebook contain codes for reproducing Fig.2 and related supplement tables of Ko et al. (https://www.biorxiv.org/content/10.1101/2022.06.13.495861v3).
Data deposited on dryad: https://doi.org/10.5061/dryad.5hqbzkh8c

### Fig.2A
```{r}
library(tidyverse) # v.2.0.0
library(RColorBrewer) # 1.1-3

## load data
path =  paste0(getwd(), "/Data/DiffExpression.tsv")
Expr = read_tsv(path) %>% 
  filter(power >= 0.8)
Expr$contrast = factor(Expr$contrast, levels = c("T1h", "T3h", "T8h", "T3d", "T7d", "T14d"))

## prep data
Expr_TF = Expr %>% 
  mutate(direction = ifelse(logFC_ave>0, "up", 'down')) %>% 
  group_by(contrast, direction, TF) %>% 
  tally()
Expr_TF$direction = factor(Expr_TF$direction, levels = c("up", "down"))

Expr_TF_lab = Expr %>% 
  mutate(direction = ifelse(logFC_ave>0, "up", 'down')) %>% 
  group_by(contrast, direction) %>% 
  tally() %>% 
  pivot_wider(id_cols = contrast, names_from = direction, values_from = n) %>% 
  mutate(y_up = down + up/2,
         y_down = down/2)
  
## pick colors
cbPalette_ori = brewer.pal(n = 3, name = "PuOr")
color_up = cbPalette_ori[3]
color_dw = cbPalette_ori[1]
plot(1:length(cbPalette_ori), rep(1, length(cbPalette_ori)), col = cbPalette_ori)
cbPalette = c(alpha(color_dw, alpha = .5), color_dw, # dw, dw_TF
              alpha(color_up, alpha = .5), color_up) # up, up_TF

FontSize = 2
p = ggplot(Expr_TF, aes(x = contrast, y = n)) + 
  geom_bar(aes(fill = interaction(TF, direction), color = direction), stat="identity", 
           position=position_stack(reverse = F), width=.75, size = .3) +
  geom_text(data = Expr_TF_lab, 
            aes(x = contrast, y = y_up, label = up), 
            color = "black", size = FontSize) +
  geom_text(data = Expr_TF_lab, 
            aes(x = contrast, y = y_down, label = down), 
            color = "black", size = FontSize) +
  scale_fill_manual(values = cbPalette) + 
  scale_color_manual(values=c(color_dw, color_up)) +
  scale_y_continuous("Number of differentially expressed genes", 
                     expand = c(0,0,0.05,0)) +
  guides(fill = guide_legend(reverse = FALSE, nrow = 4, byrow = TRUE), colour = "none") + # reverse legend
  theme_classic() + 
  theme(strip.text.y = element_blank(), 
        legend.position = 'right',
        legend.title = element_blank()
  ); p

```

### Fig.2B and Supplementary Table 5

```{r}
library(tidyverse) # v.2.0.0

## load data
path =  paste0(getwd(), "/Data/DiffExpression.tsv")
Expr = read_tsv(path) %>% 
  filter(power >= 0.8)
Expr$contrast = factor(Expr$contrast, levels = c("T1h", "T3h", "T8h", "T3d", "T7d", "T14d"))

# make wide
Expr_s = Expr %>% pivot_wider(id_cols = c(GeneSymbol), 
                              names_from = contrast, values_from = `logFC_ave`) 
Expr_s$Category = apply(Expr_s[, 2:7], 1, function(s){
  if(any(is.na(s))){"transiently regulated"}else if(all(s>0)){"constantly up-regulated"
  }else if (all(s<0)){"constantly down-regulated"}else{"dynamically regulated"}
})

```

#### plot Fig.2B

```{r}
# add count
Expr_s = Expr_s %>%
  group_by(Category) %>%
  mutate(count = n(),
         Category = paste0(Category, ": ", count))

# pick colors
cbPalette_ori = brewer.pal(n = 3, name = "PuOr")
color_up = cbPalette_ori[1] 
color_dw = cbPalette_ori[3]
plot(1:length(cbPalette_ori), rep(1, length(cbPalette_ori)), col = cbPalette_ori)

Expr_s$Color = apply(Expr_s[, 2:7], 1, function(s){
  if(any(is.na(s))){"lightgrey"}else if(all(s>0)){color_up}else if (all(s<0)){color_dw}else{"seagreen4"}
})
table(Expr_s$Color)


Expr_s_l = gather(Expr_s, contrast, LFC, T1h:T14d) %>% filter(! is.na(LFC))

order_category = sapply(c("constantly up-regulated", "constantly down-regulated", 
                          "dynamically regulated", "transiently regulated"), 
                        function(s){ unique(Expr_s$Category)[grep(s, unique(Expr_s$Category) )]})
Expr_s_l = within(Expr_s_l, Category <- factor(Category, levels = unique(order_category)))
Expr_s_l = within(Expr_s_l, contrast <- factor(contrast, levels = c("T1h", "T3h", "T8h", "T3d", "T7d", "T14d")))

range(Expr_s_l$LFC)
levels(Expr_s_l$Category)


Line_p2 = ggplot(Expr_s_l, aes(x = contrast, y = LFC, group = GeneSymbol, color = Category)) +
  geom_hline(yintercept = -0.5, size = 0.2, linetype = "dashed") +
  geom_hline(yintercept = 0.5, size = 0.2, linetype = "dashed") +
  geom_line(aes(alpha = Category)) +
  scale_color_manual(values = c(color_up, color_dw, "seagreen4", "lightgrey")) +
  scale_alpha_manual(values = c(0.2, 0.2, 0.2, 0.1)) +
  scale_y_continuous(name = expression('log'['2']*' fold change') , limits = c(-3, 3.5), 
                     breaks = c(-3, -2, -1, -0.5, 0.5, 1, 2, 3)) +
  facet_wrap(.~Category, nrow = 2) +
  # guides(alpha = 1) +
  guides(alpha = guide_legend(override.aes = list(alpha = 1)),
         color = guide_legend(nrow = 2, byrow = TRUE)
  ) + # reverse legend
  theme_classic() +
  theme(legend.position = 'none',
        panel.border = element_rect(color = 'grey50', fill = 'transparent'),
        axis.text.y = element_text(hjust = 1)) ; Line_p2

```


#### Supplementary Table 5
```{r}
Expr_s_filtered = Expr_s %>% filter(Category %in% c("constantly up-regulated", "constantly down-regulated", "dynamically regulated")) 
```


### Fig. 2D, 2E and Supplementary Table 2 and 3
```{r}
library(tidyverse) # v.2.0.0
library(lme4) # 1.1-29
library(arm) # 1.7-19
library(gridExtra) # 2.3
library(grid) # 4.1.2
set.seed(100)

## load data
path =  paste0(getwd(), "/Data/RNAscope_NormalizedStainedArea.tsv")
RNAscope = read_tsv(path)

probes = c("ALDH1A2", "AR", "HSD17B12", "SRD5A2", "UTS2B")
stats2 = lapply(probes, function(Probe_x){
  # Probe_x = probes[4]
  print(Probe_x)
  db = RNAscope %>% filter(Probe == Probe_x) 
  
  head(db)
  
  db$Group = factor(db$Group)
  db$Individual = factor(db$Individual)
  db$Slide = factor(db$Slide)
  
  db$Group <- droplevels(db$Group)
  db$Individual <- droplevels(db$Individual)
  db$Slide = droplevels(db$Slide)
  
  y_lab = "Normalized stained area (%)"
  
  # 1. deciding using sqrt or not and evaluating model
  mod_a <- lmer(data = db, sqrt(percent_area_slide) ~ Group + (1|Slide) ) # sqrt
  
  plot(mod_a)  
  summary(mod_a)
  
  plot.new()
  par(mfrow=c(3,2))
  qqnorm(resid(mod_a))
  qqline(resid(mod_a))
  qqnorm(ranef(mod_a)$"Slide"[,1])   # random effect
  qqline(ranef(mod_a)$"Slide"[,1])
  scatter.smooth(fitted(mod_a), resid(mod_a))
  
  new_data_mod_a <- expand.grid(Group = levels(db$Group)) # fix effect
  xmat <- model.matrix(~ Group, data=new_data_mod_a)  # fix effect
  new_data_mod_a$fit <- xmat%*%fixef(mod_a)
  
  nsim <- 10000
  bsim <- arm::sim(mod_a, n.sim=nsim ) #simulation of the model 10000 times
  ## To get the mean estimate of the fixed effects parameters apply(bsim@fixef, 2, mean)
  apply(bsim@fixef, 2, mean)
  ## To get the 95% credible Interval estimate of the fixed effects parameters apply(bsim@fixef, 2, quantile, prob=c(0.025, 0.975))
  apply(bsim@fixef, 2, quantile, prob=c(0.025, 0.975))
  
  Ind_b_CrI = apply(bsim@fixef, 2, quantile, prob=c(0.025, 0.975)) 
  
  Ind_b = mod_a@beta
  x = as.data.frame(VarCorr(mod_a)); x
  Ind_random_Slide = x[1, 4]
  # Ind_random_Group = x[2, 4]
  Ind_residual = x[2, 4]
  Ind_random_Slide_CrI = quantile(apply(bsim@ranef$Slide[,,1],1,var),prob=c(0.025, 0.975))
  # Ind_random_Group_CrI = quantile(apply(bsim@ranef$n_Group[,,1],1,var),prob=c(0.025, 0.975))
  Ind_residual_CrI = quantile(bsim@sigma^2,c(0.025, 0.975))
  
  df_t = rbind(cbind("Var" = Probe_x, "Parameter" = levels(db$Group), 
                     "Estimate" = Ind_b, t(Ind_b_CrI)),
               cbind("Var" = Probe_x, "Parameter" = "random_Slide", 
                     "Estimate" =  Ind_random_Slide, t(Ind_random_Slide_CrI)),
               cbind("Var" = Probe_x, "Parameter" = "residual", 
                     "Estimate" =  Ind_residual, t(Ind_residual_CrI)))
  
  fitmat <- matrix(ncol=nsim, nrow=nrow(new_data_mod_a))
  for(i in 1:nsim) fitmat[,i] <- xmat %*% bsim@fixef[i,]    # fitted values
  new_data_mod_a$lower <- apply(fitmat, 1, quantile, prob=0.025)
  new_data_mod_a$upper <- apply(fitmat, 1, quantile, prob=0.975)
  new_data_mod_a$fit <- apply(fitmat, 1, mean)
  new_data_mod_a$fit_sim<- (new_data_mod_a$fit)^2         # transform back from sqrt
  new_data_mod_a$upper_sim <- (new_data_mod_a$upper)^2   # transform back from sqrt
  new_data_mod_a$lower_sim <- (new_data_mod_a$lower)^2    # transform back from sqrt
  
  w = 0.25
  grey_point_size = 0.5
  mean_point_size = 1
  errorbar_wd = 1
  errorbar_hori = 0
  dodge = position_jitter(width = 0.05, height = 0)
  
  
  errorbar_col = "darkorange"
  
  plot_mod_a3 = ggplot(new_data_mod_a) +
    # need this layer to set up
    geom_point(data = new_data_mod_a,
               aes(x = Group, y = fit_sim), color='black', size=mean_point_size) +
    # raw data
    geom_point(data = db,
               aes(x = as.numeric(Group)+w, y = percent_area_slide),
               position = dodge, alpha = 0.5, size = grey_point_size,
               color = "grey50") +
    # CI
    geom_errorbar(data = new_data_mod_a,
                  aes(x = Group, ymax = upper_sim, ymin = lower_sim),
                  width = errorbar_hori, size = errorbar_wd,
                  color = errorbar_col) +
    # estimate
    geom_point(data = new_data_mod_a,
               aes(x = Group, y = fit_sim), color='black', size=mean_point_size) +
    ylab(y_lab) +
    ggtitle(Probe_x) +
    theme_classic() +
    theme(plot.title = element_text(hjust = 0.5)); plot_mod_a3
  
  pp = function(StimulatedTable){
    Dayn = 1:nrow(fitmat) # (row of fitmat)
    
    combi = combn(Dayn, 2)
    
    d = lapply(1:ncol(combi), function(combiN){
      # combiN = 1
      tmp = combi[, combiN]
      p = mean(fitmat[tmp[1],]<fitmat[tmp[2],])
      df = tibble("Group1" = (new_data_mod_a$Group[tmp[1]]), 
                  "Group2" = (new_data_mod_a$Group[tmp[2]]), "pp" = p)
    })
    
    dd = d %>% bind_rows() %>%
      mutate(`sig (2>1)` = ifelse(pp > 0.95, 'sig', NA))
    return(dd)
  }
  
  CI2 = Ind_b_CrI %>% t() %>% as.data.frame() %>% rownames_to_column
  df_t2 = df_t %>% as_tibble()
  
  return(list(plot_mod_a3, pp(new_data_mod_a), df_t2, CI2))
})
```
#### Supplementary Table 2 and 3
##### CredibleInterval
```{r}
pp_CI2 = lapply(1:length(stats2), function(i){
  # i=1
  stats2[[i]][[4]] %>% bind_cols("Probe" = probes[i])
}) %>% bind_rows(); pp_CI2
```

##### estimates
```{r}
pp_est2 = lapply(1:length(stats2), function(i){
  # i=1
  stats2[[i]][[3]] %>% bind_cols("Probe" = probes[i])
}) %>% bind_rows(); pp_est2
```

##### PostProbability
```{r}
pp_summary2 = lapply(1:length(stats2), function(i){
  # i=1
  stats2[[i]][[2]] %>% bind_cols("Probe" = probes[i])
}) %>% bind_rows(); pp_summary2
```

#### Fig. 2F
```{r}
plots2 =  lapply(1:length(stats2), function(i){
  # i=1
  p = stats2[[i]][[1]] #
  p
}) 

x = arrangeGrob(grobs = plots2, nrow = 1)
plot.new()
grid.draw(x)
```



### Fig. 2 - Figure supplement 1A
```{r}
library(tidyverse) # v.2.0.0

## load data
path =  paste0(getwd(), "/Data/DiffExpression.tsv")
Expr = read_tsv(path) 
Expr$contrast = factor(Expr$contrast, levels = c("T1h", "T3h", "T8h", "T3d", "T7d", "T14d"))

## prep data
Expr_TF = Expr %>% 
  mutate(direction = ifelse(logFC_ave>0, "up", 'down')) %>% 
  group_by(contrast, direction, TF) %>% 
  tally()
Expr_TF$direction = factor(Expr_TF$direction, levels = c("up", "down"))

Expr_TF_lab = Expr %>% 
  mutate(direction = ifelse(logFC_ave>0, "up", 'down')) %>% 
  group_by(contrast, direction) %>% 
  tally() %>% 
  pivot_wider(id_cols = contrast, names_from = direction, values_from = n) %>% 
  mutate(y_up = down + up/2,
         y_down = down/2)
  
## pick colors
cbPalette_ori = brewer.pal(n = 3, name = "PuOr")
color_up = cbPalette_ori[3]
color_dw = cbPalette_ori[1]
plot(1:length(cbPalette_ori), rep(1, length(cbPalette_ori)), col = cbPalette_ori)
cbPalette = c(alpha(color_dw, alpha = .5), color_dw, # dw, dw_TF
              alpha(color_up, alpha = .5), color_up) # up, up_TF

FontSize = 2
p = ggplot(Expr_TF, aes(x = contrast, y = n)) + 
  geom_bar(aes(fill = interaction(TF, direction), color = direction), stat="identity", 
           position=position_stack(reverse = F), width=.75, size = .3) +
  geom_text(data = Expr_TF_lab, 
            aes(x = contrast, y = y_up, label = up), 
            color = "black", size = FontSize) +
  geom_text(data = Expr_TF_lab, 
            aes(x = contrast, y = y_down, label = down), 
            color = "black", size = FontSize) +
  scale_fill_manual(values = cbPalette) + 
  scale_color_manual(values=c(color_dw, color_up)) +
  scale_y_continuous("Number of differentially expressed genes", 
                     expand = c(0,0,0.05,0)) +
  guides(fill = guide_legend(reverse = FALSE, nrow = 4, byrow = TRUE), colour = "none") + # reverse legend
  theme_classic() + 
  theme(strip.text.y = element_blank(), 
        legend.position = 'right',
        legend.title = element_blank()
  ); p

```


### Fig. 2 - Figure supplement 1B

```{r}
library(tidyverse) # v.2.0.0

## load data
path =  paste0(getwd(), "/Data/DiffExpression.tsv")
Expr = read_tsv(path) 
Expr$contrast = factor(Expr$contrast, levels = c("T1h", "T3h", "T8h", "T3d", "T7d", "T14d"))

# make wide
Expr_s = Expr %>% pivot_wider(id_cols = c(GeneSymbol), 
                              names_from = contrast, values_from = `logFC_ave`) 
Expr_s$Category = apply(Expr_s[, 2:7], 1, function(s){
  if(any(is.na(s))){"transiently regulated"}else if(all(s>0)){"constantly up-regulated"
  }else if (all(s<0)){"constantly down-regulated"}else{"dynamically regulated"}
})

# add count
Expr_s = Expr_s %>%
  group_by(Category) %>%
  mutate(count = n(),
         Category = paste0(Category, ": ", count))

# pick colors
cbPalette_ori = brewer.pal(n = 3, name = "PuOr")
color_up = cbPalette_ori[1] 
color_dw = cbPalette_ori[3]
plot(1:length(cbPalette_ori), rep(1, length(cbPalette_ori)), col = cbPalette_ori)

Expr_s$Color = apply(Expr_s[, 2:7], 1, function(s){
  if(any(is.na(s))){"lightgrey"}else if(all(s>0)){color_up}else if (all(s<0)){color_dw}else{"seagreen4"}
})
table(Expr_s$Color)


Expr_s_l = gather(Expr_s, contrast, LFC, T1h:T14d) %>% filter(! is.na(LFC))

order_category = sapply(c("constantly up-regulated", "constantly down-regulated", 
                          "dynamically regulated", "transiently regulated"), 
                        function(s){ unique(Expr_s$Category)[grep(s, unique(Expr_s$Category) )]})
Expr_s_l = within(Expr_s_l, Category <- factor(Category, levels = unique(order_category)))
Expr_s_l = within(Expr_s_l, contrast <- factor(contrast, levels = c("T1h", "T3h", "T8h", "T3d", "T7d", "T14d")))

range(Expr_s_l$LFC)
levels(Expr_s_l$Category)


Line_p2 = ggplot(Expr_s_l, aes(x = contrast, y = LFC, group = GeneSymbol, color = Category)) +
  geom_hline(yintercept = -0.5, size = 0.2, linetype = "dashed") +
  geom_hline(yintercept = 0.5, size = 0.2, linetype = "dashed") +
  geom_line(aes(alpha = Category)) +
  scale_color_manual(values = c(color_up, color_dw, "seagreen4", "lightgrey")) +
  scale_alpha_manual(values = c(0.2, 0.2, 0.2, 0.1)) +
  scale_y_continuous(name = expression('log'['2']*' fold change') , limits = c(-3, 3.5), 
                     breaks = c(-3, -2, -1, -0.5, 0.5, 1, 2, 3)) +
  facet_wrap(.~Category, nrow = 2) +
  # guides(alpha = 1) +
  guides(alpha = guide_legend(override.aes = list(alpha = 1)),
         color = guide_legend(nrow = 2, byrow = TRUE)
  ) + # reverse legend
  theme_classic() +
  theme(legend.position = 'none',
        panel.border = element_rect(color = 'grey50', fill = 'transparent'),
        axis.text.y = element_text(hjust = 1)) ; Line_p2

```

### Fig. 2 - Figure supplement 1C
```{r}
library(tidyverse) # v.2.0.0

## load data
path =  paste0(getwd(), "/Data/DiffExpression.tsv")
Expr = read_tsv(path) 
Expr$contrast = factor(Expr$contrast, levels = c("T1h", "T3h", "T8h", "T3d", "T7d", "T14d"))

Expr2 = Expr %>% dplyr::select(GeneSymbol, logFC_ave, contrast) %>% distinct() %>% 
  pivot_wider(id_cols = c(GeneSymbol), 
              names_from = contrast, values_from = logFC_ave) 
Expr2$`Number of regulated time points` =  apply(Expr2[, 2:7], 1, function(s){
  sum(!is.na(s))
})

Expr2_sum = Expr2 %>% group_by(`Number of regulated time points`) %>% 
  tally(name = "Number of genes") %>% 
  mutate(class = ifelse(`Number of regulated time points` == 6, 'constantly regulated', 'transiently regulated'))
Expr2_sum$class <- factor(Expr2_sum$class, levels = c('transiently regulated', 'constantly regulated'))

p = ggplot(Expr2_sum, aes(x = `Number of regulated time points`, y = `Number of genes`)) +
  geom_col(aes(fill = class)) +
  geom_text(aes(label = `Number of genes`), vjust = -1, size = FontSize) +
  scale_y_continuous(expand = c(0,0,0.05,0)) + 
  scale_x_continuous(breaks = 1:6) + 
  scale_fill_manual(values = c('grey70', 'grey30')) +
  theme_classic() +
  theme(legend.position = c(0.7, 0.85)); p

```

### Fig. 2 - Figure supplement 1D
```{r}
library(tidyverse) # v.2.0.0

## load data
path =  paste0(getwd(), "/Data/DiffExpression.tsv")
Expr = read_tsv(path) %>% 
  filter(power >= 0.8)
Expr$contrast = factor(Expr$contrast, levels = c("T1h", "T3h", "T8h", "T3d", "T7d", "T14d"))

Expr2 = Expr %>% dplyr::select(GeneSymbol, logFC_ave, contrast) %>% distinct() %>% 
  pivot_wider(id_cols = c(GeneSymbol), 
              names_from = contrast, values_from = logFC_ave) 
Expr2$`Number of regulated time points` =  apply(Expr2[, 2:7], 1, function(s){
  sum(!is.na(s))
})

Expr2_sum = Expr2 %>% group_by(`Number of regulated time points`) %>% 
  tally(name = "Number of genes") %>% 
  mutate(class = ifelse(`Number of regulated time points` == 6, 'constantly regulated', 'transiently regulated'))
Expr2_sum$class <- factor(Expr2_sum$class, levels = c('transiently regulated', 'constantly regulated'))

p = ggplot(Expr2_sum, aes(x = `Number of regulated time points`, y = `Number of genes`)) +
  geom_col(aes(fill = class)) +
  geom_text(aes(label = `Number of genes`), vjust = -1, size = FontSize) +
  scale_y_continuous(expand = c(0,0,0.05,0)) + 
  scale_x_continuous(breaks = 1:6) + 
  scale_fill_manual(values = c('grey70', 'grey30')) +
  theme_classic() +
  theme(legend.position = c(0.7, 0.85)); p

```
